A Study on Common Bug Fix Patterns in Rust

### 2022-10-27 sections/methodology.tex
encode to embed

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints. Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.

@@

An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit. Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and encoding that change in our datapoints. Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.

@@

Figure~\ref{fig:extraction} illustrates path extraction on a simple change. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint encoding.

\paragraph{Encoding the data.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}. The number of columns would depend on the amount of information we want to encode in the datapoints.

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an encoding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to encode all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive encodings still holds.

@@

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description is the most useful one, and we designed our weighting scheme to encode this value judgment.

>>>>

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and embed the important parts of the derived AST in fixed-size datapoints. Work in this vein makes the assumption that each cluster corresponds to one change pattern~\cite{hanam2016discovering,campos2019discovering,yang2022mining}.

@@

An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit. Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and embedding that change in our datapoints. Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.

@@

Figure~\ref{fig:extraction} illustrates path extraction on a simple change. The change replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. That is, after the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, in the third argument, instead of sending \verb+head+ to the callee, it sends a clone of it. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. We found it crucial to record the order of nodes in the path, as this gives us the freedom to later record the order in our datapoint embedding.

\paragraph{Code Embedding.} A key question is how to transform these paths to a fixed size datapoint. First, we need to define a fixed set of columns. We need fixed-size datapoints, as they are a requirement for both centroid-based and density-based well-known clustering algorithms~\cite{xu2005survey}. The number of columns would depend on the amount of information we want to embed in the datapoints.

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields an infeasible $O(n^d)$ dimensions ($n=|\mathtt{NT} \cup \mathtt{BCE}|$ and $d$ is the depth of nested structures); this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.

@@

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description is the most useful one, and we designed our weighting scheme to embed this value judgment.

### 2022-10-27 sections/results.tex
fixed rq1 numbers
encode to embed

As discussed in the previous section, we used an encoding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (weighting scheme) while encoding our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements). Here we want to evaluate the effectiveness of our encoding.

The general idea for our evaluation is whether the most important program elements get the high values in our encoding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point~\cite{collins2003circle}. For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our encoding (similar to Figure~\ref{fig:essence}).

@@

We carried out two manual experiments for evaluating our encoding mechanism. In the first experiment, the analyzer investigates whether the most important element has the largest radius or not. In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not. The analyzer should give a binary answer to these questions. Both authors carried out these experiments independently. At last, we report the results as the portion of the changes which both authors agreed upon its encoding correctness. The final results are 55\% and 78\% correctness in the first and second experiments, respectively, which shows the effectiveness of our encoding mechanism.

>>>>

As discussed in the previous section, we used an embedding mechanism to incorporate AST information in fixed-sized datapoints which then could be used in the DBSCAN clustering algorithm. We used heuristics (weighting scheme) while creating our datapoints to make the most important program elements stand out between all the non terminals seen in the AST (alongside some BC-related elements). Here we want to evaluate the effectiveness of our code embedding.

The general idea for our evaluation is whether the most important program elements get the high values in our embedding. The analyzer compares the actual change in the code and compares it to the visual representation of its respective data point. We used circle pack figures to visualize each data point~\cite{collins2003circle}. For this purpose, we used the packcircles python library\footnote{\url{https://github.com/mhtchan/packcircles}}. In these figures, the radius of the circle of each non terminal shows its importance in our embedding (similar to Figure~\ref{fig:essence}).

@@

We carried out two manual experiments for evaluating our embedding mechanism. In the first experiment, the analyzer investigates whether the most important element has the largest radius or not. In the second experiment, the analyzer investigates whether the most important element is included in three largest circles or not. The analyzer should give a binary answer to these questions. Both authors carried out these experiments independently. At last, we report the results as the portion of the changes which both authors agreed upon its embedding correctness. The final results are 54\% and 80\% correctness in the first and second experiments, respectively.
